### 2020-03-06 short.body.tex
edits

In an epidemic, symptomatic cases are the predominant focus of treatment and usually represent the bulk of reported cases. 
However, infected individuals who are asymptomatic yet infectious can be a critical factor in the spread of some pathogens.
Asymptomatic individuals are hard to trace, unlikely to self-isolate, and are likely to retain normal social and travel patterns.
The COVID-19 outbreak has raised significant
questions regarding the role of asymptomatic cases~\citep{fauci_nejm2020}.
A key focus has been on estimating asymptomatic prevalence (see~\citep{mizumoto_2020}), e.g., what fraction of individuals have asymptomatic infections relative to those with symptomatic infections?
%.(i) What fraction of newly infected individuals
%have mild and/or subclinical infections rather than severe and/or
%clinical infections? (ii) What is the time-scale over which 
%individuals who have mild and/or subclinical cases remain
%infectious? (iii) How infectious are individuals
%with mild and/or subclinical infections relative to those
%with severe symptoms?

There is currently a lot of interest in the role of asymptomatic infections in COVID for two major reasons. 
First, the proportion of infections that are asymptomatic is critical to attempts to estimate the likely burden of severe outcomes (including mortality) when the virus spreads through a population.
Second, understanding the possible role of \emph{transmission} by asymptomatic individuals is crucial to planning surveillance and control efforts.

@@

We model viral spread using a renewal-equation framework, with incidence $i$ is divided into two categories -- $i_a$ and $i_s$ -- corresponding to incidence of asymptomatic and symptomatic cases, respectively. 
Both types of individuals can infect others, but they differ in their intrinsic reproduction numbers, ${\cal{R}}_a$ and ${\cal{R}}_s$, as well as intrinsic generation-interval distributions, $g_a(\tau)$ and $g_s(\tau)$.
Generation intervals, which are defined as the time between when an individual is infected and when that individual infects another person \citep{svensson2007note}, depend on the natural history of infection:
individuals with subclinical cases may have fast clearance and short generation intervals, or slow viral reproduction and long generation intervals.

@@

The basic reproduction number of this system \citep{anderson1992infectious} is:
\begin{equation}
{\cal{R}}_{0}= p {\cal{R}}_a + (1-p) {\cal{R}}_s.
\end{equation}
We define the ``intrinsic'' proportion of asymptomatic transmission as the relative contribution of asymptomatic cases to the basic reproduction number:
\begin{equation}
z = p {\cal{R}}_a/{\cal{R}}_{0}
\end{equation}
Note that the intrinsic proportion of symptomatic transmission satisfies
\begin{equation}
1-z = (1-p) {\cal{R}}_s/{\cal{R}}_{0}.
\end{equation}

@@

Here, given $1/r=7$ days, $\bar G_s=6$ days, and $\kappa_s=0.5$, we evaluate the impacts of intrinsic proportion of asymptomatic transmission $z$ on the realized relevance of asymptomatic infection, given variation in relative mean generation interval $\bar G_a/\bar G_s$ and relative generation-interval dispersion $\kappa_a/\kappa_s$ of asymptomatic cases.
We vary $\bar G_a/\bar G_s$ and $\kappa_a/\kappa_s$ between 0.5 and 2 one at a time, and infer $q$ and ${\cal{R}}_0$ compatible with an estimated value of $r$ from case data given the generation interval distributions of asymptomatic and asymptomatic cases.
Across the ranges of parameters we explore, the intrinsic proportion of asymptomatic transmission $z$ is similar to the realized proportion $q$ (Figure~\ref{fig.relevance}A,C).
When the mean asymptomatic generation interval is longer than the mean symptomatic generation interval, $q$ decreases because short generation intervals (i.e., fast transmission) caused by symptomatic individuals drive the transmission during the growth phase (Figure~\ref{fig.relevance}A).
On the other hand, assuming a wide generation-interval distribution (higher $\kappa_a$) for asymptomatic individuals increases $q$ because the amount of short generation intervals (i.e., the proportion of faster transmission by asymptomatic individuals) increases (Figure~\ref{fig.relevance}C).

Figure~\ref{fig.relevance}C shows that when $\bar G_s<\bar G_a$, then the strength ${\cal{R}}_0$ increases with $z$.
This increase arises because asymptomatic infections that have long generation intervals increase the duration of overall generation intervals (see~\citep{park_2019practical})
Conversely, when $\bar G_a > \bar G_s$ then the strength ${\cal{R}}_0$ decreases with $z$ because short generation intervals of asymptomatic cases reduce the overall mean generation interval.
Likewise, the amount of dispersion in the generation intervals of asymptomatic cases $\kappa_a$ affect the amount of dispersion in the overall generation intervals; 
therefore, depending on whether $\kappa_a$ is larger or smaller than $\kappa_s$, increase in the proportion of asymptomatic transmission $z$ can either decrease or increase the strength ${\cal{R}}_0$ (Figure~\ref{fig.relevance}D).
In other words, when ${\cal{R}}_0$ is estimated without explicitly accounting for asymptomatic spread, it can be over- or under- estimated depending on the relative mean and dispersion in generation intervals of asymptomatic and symptomatic individuals.

@@

The present findings highlight the need to characterize
the course of infection of asymptomatic cases and their consequences
not only for contact tracing but for constraining
uncertainty in the strength of the ongoing COVID-19 outbreak~\citep{park_preprint}.
The present findings are also consistent with
a recent generalization linking speed, strength, and
generation intervals, such that
for a given observed speed
increases in the mean generation interval imply
larger reproduction number~\citep{park_2019practical}. Hence, reports of individuals
with mild, persistent cases 
suggest the possibility that the mean generation interval for COVID-19
could be longer than estimated from symptomatic cases alone - possibly
increasing ${\cal{R}}_0$ and driving a larger fraction of secondary cases
(see Figure~\ref{fig.relevance} top row).
However, if asymptomatic cases tend to have shorter infection, then
current estimates of ${\cal{R}}_0$ may be over-estimates of the
underlying strength (see Figure~\ref{fig.relevance} bottom row).
Finally, it is important to note we focused here on the exponential
phase, but that the asymptomatic relevance $q$ is time-dependent, and will
vary given the population state structure of infected individuals.  Such
variation in the ratio of asymptomatic to symptomatic cases
will also impact estimates of case fatality rates insofar
as many asymptomatic cases are not counted, particularly early in an outbreak.
Future work might also consider the ways in which asymptomatic individuals
modulate resurgent epidemics in a networked metapopulation~\citep{watts_pnas2005}.

\begin{figure*}[b!]
\begin{center}
\includegraphics[width=0.4\textwidth]{figheatmap.pdf}
\mbox{\hspace{0.05\textwidth}}
\includegraphics[width=0.4\textwidth]{figheatmap_R0.pdf}\\
\mbox{}\\
\includegraphics[width=0.4\textwidth]{figheatmap_kappa.pdf}
\mbox{\hspace{0.05\textwidth}}
\includegraphics[width=0.4\textwidth]{figheatmap_kappa_R0.pdf}
\caption{Relevance of asymptomatic transmission given variation in
probability that a given, newly infected individual is asymptomatic. (Top) Long/short generation intervals of asymptomatic cases increase/decrease ${\cal{R}}_0$ with decreasing/increasing relevance.
(Bottom) Wide/narrow generation intervals of asymptomatic cases decrease/increase ${\cal{R}}_0$ with increasing/decreasing relevance.
% Technically, renewal equations provide a route to connect
% observed increases in case counts, i.e., ${\cal{R}}_0 = \frac{1}{M(-r)}$
% where $M(-r)=\int_0^{\infty}\mathrm{d} a e^{-ra}g(a)$.
% For the present model,
% $M(r)=qM_E(-r)M_{I_a}(-r)+(1-q)M_E(-r)M_{I_s}(-r)$
% where each of the component functions are $M_E(-r)=\frac{\gamma_e}{\gamma_e+r}$,
% $M_{I_a}(-r)=\frac{\gamma_a}{\gamma_a+r}$, 
% $M_{I_s}(-r)=\frac{\gamma_s}{\gamma_s+r}$,
% and $q$ is the fraction of realized new secondary cases caused by
% asymptomatic individuals.  In all of these heatmaps, $\gamma_e=1/4$,
% $\gamma_s=1/10$, and $r=1/7$. Here, the average duration
% in symptomatic and asymptomatic states are $T_s=1/\gamma_s$ and 
% $T_a=1/\gamma_a$, respectively. Here the filled circle denotes ${\cal{R}}_0\approx 3.12$
% in the event of $p=0$ and $\beta_a=0$, and the open symbols are for $\beta_a/\beta_s=0.5$ and 
% $p=0.5$ and $p=0.9$ for reference.
%Renewal equations provide a route to connect
%observed increases in case counts, i.e.,
%\begin{equation}
%{\cal{R}}_0 = \frac{1}{M(-r)}
%\end{equation}
%where $r$ is the speed of the disease spread and 
%$M(r)$ is related to 
%$g(a)$ (the normalized age of transmission) such that 
%\begin{equation}
%M(r)=\int_0^{\infty}\mathrm{d} a e^{ra}g(a).
%\end{equation}
%For the present model, the moment generating function is
%\begin{equation}
%M(r)=qM_E(r)M_{I_a}(r)+(1-q)M_E(r)M_{I_s}(r)
%\end{equation}
%where each of the component functions are $M_E(-r)=\frac{\gamma_e}{\gamma_e+r}$,
%$M_{I_a}(-r)=\frac{\gamma_a}{\gamma_a+r}$, 
%$M_{I_s}(-r)=\frac{\gamma_s}{\gamma_s+r}$,
%and $q$ is the fraction of realized new secondary cases caused by
%asymptomatic individuals. 
%The relevance $q$ is an unknown feature of
\label{fig.relevance}}
\end{center}
\end{figure*}

>>>>

\section{Introduction}

In an epidemic, symptomatic cases are the predominant focus of treatment and usually represent the bulk of reported cases. 
However, infected individuals who are asymptomatic yet infectious can be a critical factor in the spread of some pathogens \citep{fraser2004factors}.
Asymptomatic individuals are hard to trace, unlikely to self-isolate, and are likely to retain normal social and travel patterns.
The COVID-19 outbreak has raised significant
questions regarding the role of asymptomatic cases~\citep{fauci_nejm2020}.
A key focus has been on estimating asymptomatic prevalence (see~\citep{mizumoto_2020}), e.g., what fraction of individuals have asymptomatic infections relative to those with symptomatic infections?
%.(i) What fraction of newly infected individuals
%have mild and/or subclinical infections rather than severe and/or
%clinical infections? (ii) What is the time-scale over which 
%individuals who have mild and/or subclinical cases remain
%infectious? (iii) How infectious are individuals
%with mild and/or subclinical infections relative to those
%with severe symptoms?

There is currently a lot of interest in the role of asymptomatic infections in SARS-CoV-2 for two major reasons. 
First, the proportion of infections that are asymptomatic is critical to attempts to estimate the likely burden of severe outcomes (including mortality) when the virus spreads through a population.
Second, understanding the possible role of \emph{transmission} by asymptomatic individuals is crucial to planning surveillance and control efforts.

@@

\section{Methods}

We model viral spread using a renewal-equation framework \citep{heesterbeek1996concept}, which allows us to model the current incidence of infected individuals as a function of previous incidence and how infectiousness of an infected individual varies over the course of their infection;
in Appendix, we show that results derived from a compartmental model is a special case of the result we present here.
We divide incidence $i$ into two categories -- $i_a$ and $i_s$ -- corresponding to incidence of asymptomatic and symptomatic cases, respectively.
Both types of individuals can infect others, but they differ in their intrinsic reproduction numbers, ${\cal{R}}_a$ and ${\cal{R}}_s$, as well as intrinsic generation-interval distributions, $g_a(\tau)$ and $g_s(\tau)$.
Generation intervals, which are defined as the time between when an individual is infected and when that individual infects another person \citep{svensson2007note}, depend on the natural history of infection:
individuals with subclinical cases may have fast clearance and short generation intervals, or slow viral reproduction and long generation intervals.

@@

The basic reproduction number -- the expected number of secondary cases generated by an average primary case in a fully susceptible population \citep{anderson1992infectious} -- of this system is:
\begin{equation}
{\cal{R}}_{0}= p {\cal{R}}_a + (1-p) {\cal{R}}_s.
\end{equation}
We define the ``intrinsic'' proportion of asymptomatic transmission as the relative contribution of asymptomatic cases to the basic reproduction number:
\begin{equation}
z = p {\cal{R}}_a/{\cal{R}}_{0}.
\end{equation}
Note that the intrinsic proportion of symptomatic transmission satisfies
\begin{equation}
1-z = (1-p) {\cal{R}}_s/{\cal{R}}_{0}.
\end{equation}

@@

Here, given $1/r=7$ days, $\bar G_s=7$ days, and $\kappa_s=0.5$, we evaluate the impacts of intrinsic proportion of asymptomatic transmission $z$ on the realized relevance of asymptomatic infection, given variation in relative mean generation interval $\bar G_a/\bar G_s$ and relative generation-interval dispersion $\kappa_a/\kappa_s$ of asymptomatic cases.
We vary $\bar G_a/\bar G_s$ and $\kappa_a/\kappa_s$ between 0.5 and 2 one at a time, and infer $q$ and ${\cal{R}}_0$ compatible with an estimated value of $r$ from case data given the generation interval distributions of asymptomatic and asymptomatic cases.

\section{Results}

Across the ranges of parameters we explore, the intrinsic proportion of asymptomatic transmission $z$ is similar to the realized proportion $q$ (Figure~\ref{fig.relevance}A,C).
As the relative mean generation interval of asymptomatic cases, $\bar G_a/\bar G_s$, increases, $q$ decreases because symptomatic cases are more likely to have short generation intervals (i.e., fast transmission events), which drive the spread during the growth phase (Figure~\ref{fig.relevance}A).
On the other hand, as the relative generation-interval dispersion of asymptomatic cases, $\kappa_a/\kappa_s$, increases, $q$ also increases because the amount of short asymptomatic generation intervals (i.e., the proportion of faster transmission by asymptomatic individuals) increases (Figure~\ref{fig.relevance}C).

Figure~\ref{fig.relevance}B compares the effect of relative mean generation interval of asymptomatic cases, $\bar G_a/\bar G_s$, on ${\cal{R}}_0$.
When $\bar G_s < \bar G_a$, then the strength ${\cal{R}}_0$ increases with $z$.
This increase arises because asymptomatic infections that have long generation intervals increase the duration of overall generation intervals (see~\citep{park_2019practical})
Conversely, when $\bar G_s > \bar G_a$ then the strength ${\cal{R}}_0$ decreases with $z$ because short generation intervals of asymptomatic cases reduce the overall mean generation interval.

Likewise, Figure~\ref{fig.relevance}D compares the effect of relative generation-interval dispersion of asymptomatic cases, $\kappa_a/\kappa_s$, on ${\cal{R}}_0$.
Depending on whether $\kappa_a$ is larger or smaller than $\kappa_s$, increase in the proportion of asymptomatic transmission $z$ can either decrease or increase the strength ${\cal{R}}_0$.
In other words, when ${\cal{R}}_0$ is estimated without explicitly accounting for asymptomatic spread, it can be over- or under- estimated depending on the relative mean and dispersion in generation intervals of asymptomatic and symptomatic individuals.

@@

\section{Discussion}

The present findings highlight the need to characterize the course of infection of asymptomatic cases and their consequences not only for contact tracing but for constraining uncertainty in the strength of the ongoing COVID-19 outbreak~\citep{park_preprint}.
The present findings are also consistent with a recent generalization linking speed, strength, and generation intervals, such that for a given observed speed increases in the mean generation interval imply larger reproduction number~\citep{park_2019practical}. 
Hence, reports of individuals with mild, persistent cases suggest the possibility that the mean generation interval for COVID-19 could be longer than estimated from symptomatic cases alone - possibly increasing ${\cal{R}}_0$ (Figure~\ref{fig.relevance}B).
However, if asymptomatic cases tend to have shorter infection, then current estimates of ${\cal{R}}_0$ may be over-estimates of the underlying strength (Figure~\ref{fig.relevance}B), although asymptomatic cases may be driving a larger fraction of secondary cases than we would expect when both asymptomatic and symptomatic cases have identical generation-interval distributions (Figure~\ref{fig.relevance}A).

Finally, it is important to note we focused here on the exponential phase, but that the asymptomatic relevance $q$ is time-dependent, and will vary given the population state structure of infected individuals. Such variation in the ratio of asymptomatic to symptomatic cases will also impact estimates of case fatality rates insofar as many asymptomatic cases are not counted, particularly early in an outbreak.
Future work might also consider the ways in which asymptomatic individuals modulate resurgent epidemics in a networked metapopulation~\citep{watts_pnas2005}.
Characterizing the role of asymptomatic individuals in driving the persistence of the epidemic will be critical for assessing the post-pandemic outcome \citep{lipsitch_preprint}.

\begin{figure*}[b!]
\begin{center}
\includegraphics[width=0.4\textwidth]{figheatmap.pdf}
\mbox{\hspace{0.05\textwidth}}
\includegraphics[width=0.4\textwidth]{figheatmap_R0.pdf}\\
\mbox{}\\
\includegraphics[width=0.4\textwidth]{figheatmap_kappa.pdf}
\mbox{\hspace{0.05\textwidth}}
\includegraphics[width=0.4\textwidth]{figheatmap_kappa_R0.pdf}
\caption{Relevance of asymptomatic transmission given variation in
probability that a given, newly infected individual is asymptomatic. 
(A) Long/short generation intervals of asymptomatic cases decrease/increase the relevance of asymptomatic cases, $q$.
(B) Long/short generation intervals of asymptomatic cases increase/decrease the basic reproduction number ${\cal{R}}_0$.
(C) Wide/narrow generation intervals of asymptomatic cases increase/decrease the relevance of asymptomatic cases, $q$.
(D) Wide/narrow generation intervals of asymptomatic cases decrease/increase the basic reproduction number ${\cal{R}}_0$.
For panels A--B, we vary relative mean generation interval of asymptomatic cases, $\bar G_a/\bar G_s$, while assuming $\kappa_a = \kappa_s$.
For panels C--D, we vary relative generation-interval dispersion of asymptomatic cases, $\kappa_a/\kappa_s$, while assuming $G_a = G_s$.
As our baseline parameters, we assume $1/r=7$ days, $\bar G_s=7$ days, and $\kappa_s=0.5$.
% Technically, renewal equations provide a route to connect
% observed increases in case counts, i.e., ${\cal{R}}_0 = \frac{1}{M(-r)}$
% where $M(-r)=\int_0^{\infty}\mathrm{d} a e^{-ra}g(a)$.
% For the present model,
% $M(r)=qM_E(-r)M_{I_a}(-r)+(1-q)M_E(-r)M_{I_s}(-r)$
% where each of the component functions are $M_E(-r)=\frac{\gamma_e}{\gamma_e+r}$,
% $M_{I_a}(-r)=\frac{\gamma_a}{\gamma_a+r}$, 
% $M_{I_s}(-r)=\frac{\gamma_s}{\gamma_s+r}$,
% and $q$ is the fraction of realized new secondary cases caused by
% asymptomatic individuals.  In all of these heatmaps, $\gamma_e=1/4$,
% $\gamma_s=1/10$, and $r=1/7$. Here, the average duration
% in symptomatic and asymptomatic states are $T_s=1/\gamma_s$ and 
% $T_a=1/\gamma_a$, respectively. Here the filled circle denotes ${\cal{R}}_0\approx 3.12$
% in the event of $p=0$ and $\beta_a=0$, and the open symbols are for $\beta_a/\beta_s=0.5$ and 
% $p=0.5$ and $p=0.9$ for reference.
%Renewal equations provide a route to connect
%observed increases in case counts, i.e.,
%\begin{equation}
%{\cal{R}}_0 = \frac{1}{M(-r)}
%\end{equation}
%where $r$ is the speed of the disease spread and 
%$M(r)$ is related to 
%$g(a)$ (the normalized age of transmission) such that 
%\begin{equation}
%M(r)=\int_0^{\infty}\mathrm{d} a e^{ra}g(a).
%\end{equation}
%For the present model, the moment generating function is
%\begin{equation}
%M(r)=qM_E(r)M_{I_a}(r)+(1-q)M_E(r)M_{I_s}(r)
%\end{equation}
%where each of the component functions are $M_E(-r)=\frac{\gamma_e}{\gamma_e+r}$,
%$M_{I_a}(-r)=\frac{\gamma_a}{\gamma_a+r}$, 
%$M_{I_s}(-r)=\frac{\gamma_s}{\gamma_s+r}$,
%and $q$ is the fraction of realized new secondary cases caused by
%asymptomatic individuals. 
%The relevance $q$ is an unknown feature of
\label{fig.relevance}}
\end{center}
\end{figure*}
